### 2024-11-23 paper/paper.tex

\documentclass[sigplan,11pt,nonacm]{acmart}
\settopmatter{printfolios}

\usepackage{booktabs} % For formal tables
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{hyphenat}
\usepackage{todonotes}
\usepackage[babel]{csquotes}

\begin{document}
\title{Utilizing Parallel Workers: \\LLVM's Vectorization Plan}
\author{Jonas Fritsch}
\affiliation{%
  \institution{Technical University of Munich}
}
\email{jonas.fritsch@tum.de}

\begin{abstract}
  Modern SIMD processors provide various vector registers and ISAs for programmers to utilize. 
  However, manually vectorizing code can be time-consuming, so compilers look to auto-vectorize 
  code for better performance. LLVM, as a widely used compiler toolchain, could only auto-vectorize 
  basic scalar code. Intel proposed a comprehensive refactoring of the underlying system to enable 
  LLVM's auto-vectorization to support more complex optimizations. The suggested Vectorization Plan 
  (VPlan) architecture would be more modular and scalable. Since then, significant efforts have been 
  made to implement this new system. Today, VPlan is already actively involved in vectorizing 
  innermost loops. Other areas, such as outer loop vectorization, are also in development.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}
Modern CPUs are often equipped with multiple different vector registers. These registers are nowadays 
as wide as 512 bits, allowing for the processing of multiple data streams at once (SIMD). By 
batching multiple values together in one register, different ISAs like Intel AVX or ARM SVE allow 
the execution of the same instruction for all values in a vector register simultaneously. 
Utilizing this leads to significant performance improvements over the scalar equivalent most of 
the time.

However, manual code vectorization can quickly become time-consuming, especially when
supporting different CPU architectures. Modern compilers aim to automatically transform scalar code
to use vectorization when applicable.

As one of the most widely used compilation frameworks, LLVM~\cite{10.5555/977395.977673} has 
implemented and refined its auto-vectorization over many years. It provides two different 
Vectorizers, one for innermost loops (LoopVectorize) and one for super-word parallelism 
(SLPVectorize)~\cite{llvmvec}.

This system, however, had quite a few limitations, as the loop vectorizer could only handle
innermost loops and neither outer loops, complex control flow, nor non-inlined functions. 
Additionally, while multiple different vectorizations for the same scalar code 
would be possible, the current vectorizers working directly on the LLVM IR could not model
and compare the costs of such different vectorization approaches.

With these limitations in mind, Intel started an ongoing refactorization effort to migrate LLVM's
auto-vectorization pipeline to utilize a more abstract Vectorization Plan 
(VPlan)~\cite{llvmextloopvec,llvmvplan}. The final goal would be to unite LLVM's auto-vectorization
in a single flexible system capable of optimizing SLPs, inner, and nested loops with complex 
control flows. The auto-vectorization pass would create and transform multiple different 
VPlans, each modeling a different vectorization approach, abstracted away from the underlying LLVM IR.
Those VPlans are compared against each other based on a cost model, and finally, the best one, which
might also not vectorize at all, is materialized into the LLVM IR.

LLVM's VPlan is currently being integrated into the existing Loop Vectorizer and is
already modeling most inner loop vectorizations and transformations~\cite{llvmvplanupdate}. 
Vectorization for outer loops is also in development and can be enabled by setting 
the \texttt{-enable-\allowbreak vplan-\allowbreak native-\allowbreak path} feature 
flag~\cite{llvmouterloop,llvmouterloopstatus}. 
In the future, the plan will be to merge both of these loop vectorization paths into one. 
Merging the existing SLP Vectorizer with VPlan is also planned.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background}
\label{sec:background}
In general, vectorization techniques can be separated into different categories:

\subsection{Inner Loop Vectorization}
Innermost loops are loops whose body's control flow does not contain any more loops, bound or unbound.
Vectorization of these loops is more straightforward as the only option to vectorize is over 
the loop's induction variable.

There are various ways to transform the control flow of a loop to accommodate operating on widened
vector registers. Figure~\ref{fig:inner-loop-vec} shows an example of an if-conversion. Some more
examples can also be found in LLVM's Auto-Vectorization Documentation~\cite{llvmvec}.

\begin{figure}
  \centering
  \includegraphics[width=0.35\textwidth]{images/inner-loop-vec.png}
  \caption{A code snippet of a simple loop that would be vectorized over the variable \texttt{i}. 
  The conditional branch would be flattened into a masked addition. The mask would be generated 
  based on \texttt{z[i]}.}
  \label{fig:inner-loop-vec}
\end{figure}

\subsection{Super-Word Parallelism (SLP) Vectorization}
SLP Vectorization vectorizes similar independent instructions. A compiler might have, for example, 
fully unrolled a smaller loop in a previous code simplification pass. The resulting similar scalar 
instructions might then be vectorized later by an SLP vectorizer.

\subsection{Outer Loop Vectorizaton}
Outer loop vectorization focuses on control flow with nested loops. Such loops often present
more challenges for auto-vectorization, as one could vectorize over the outer loop,
inner loop, or a mix of both, representing a hybrid/multi-dimensional vectorization approach. 
Carrying out and comparing costs of such different vectorization approaches can be complex 
and require a flexible system like, for example, VPlans.

Figure~\ref{fig:outer-loop-vec} shows a nested loop code snippet where vectorization over the
outer loop would be beneficial.

\begin{figure}
  \centering
  \includegraphics[width=0.39\textwidth]{images/outer-loop-vec.png}
  \caption{A code snippet of a nested loop where it would be beneficial to vectorize over the 
  outer loop. The inner loop has a trip count that is too low (6) to vectorize. Additionally,
  the memory access pattern of \texttt{y} would be consecutive compared to scattered when
  vectorizing over \texttt{i} instead of \texttt{j}.}
  \label{fig:outer-loop-vec}
\end{figure}

\subsection{Function Vectorization}
Function vectorization refers to the vectorization of entire functions. A loop might contain a
non-inlined function call, passing data related to the iteration variable. During auto-vectorization, 
this call could be replaced with a call to a newly constructed function with the same semantic 
control flow but operating on multiple values at once through vector registers instead.

\subsection{Vectorization Similarities}
All of the above-mentioned vectorization techniques can be brought into relation with each other. 
For example, function vectorization can be transformed into loop vectorization by creating a loop 
around the function body and vectorizing it and vice versa. The same applies for SLP Vectorization 
and loop vectorization when unrolling a loop.

\subsection{Vectorization Legality}
Not all loops can be vectorized. There are various factors that can ultimately hinder any
vectorization attempts.

Some of the most common vectorization barriers are data dependencies. As the vectorized version
of a code executes for Vectorization Factor (VF) $N$ data streams simultaneously, it inherently changes
the order of operations. To guarantee that the semantics of the code do not change, all
operations on these $N$ data streams must, generally, be independent. The simplest case
where this is not provided is when the $i$-th loop iteration depends on any of the $(i-N)$-th 
previous iterations.

While explicit data dependencies are easier to see, a simple loop, as in 
Figure~\ref{fig:inner-loop-vec}, could carry an implicit dependency in line 3 if
the \texttt{x} and \texttt{y} point to overlapping memory regions (pointer aliasing).

Not all data dependencies directly hinder any vectorization. One exception can 
sometimes be made for reduction idioms, as seen in the inner loop in 
Figure~\ref{fig:outer-loop-vec}, where \texttt{sum} would be a reduction variable.

Additionally, the control flow should be primarily straight-line code. While most if-else 
constructs can be flattened to always execute both branches along with masking, the same is not 
valid for switch-case statements or too complex control flows.

\subsection{Vectorization Costs}
While vectorizing a given control flow may not be legal, the vectorized code sometimes performs 
worse than its scalar equivalent.

Some control flow transformations that seem beneficial at compile time might reveal a performance
bottleneck at runtime. Considering the code snippet from Figure~\ref{fig:inner-loop-vec}, it might
be that almost all \texttt{z[i]} evaluate to false at runtime. The branch predictor of a CPU
would then be able to guess the correct branch in the scalar code. The vectorized version would
always load the \texttt{y[i]} into memory only for the addition to be masked away. Different 
approaches to avoid such runtime penalties are branch-on-superword-condition code 
(BOSCC)~\cite{10.5555/1299042.1299055,llvmboscc} or active-lane 
consolidation (ALC)~\cite{10.1007/s11227-022-04359-w,10.5555/3615924.3615932}.

Some vector instructions can be costly on most CPUs. Examples are gather, scatter, shuffle,
and permutation instructions. Control flow with many non-adjacent memory loads/stores or 
operations that combine different vector elements, such as reductions, can decrease vectorization 
performance.

Additionally, the vectorized code size will always be larger than the scalar equivalent. This is
due to many factors: (1) The transformed control flow often includes more instructions.
(2) As the loop trip count might not be a perfect multiple of the vectorization factor, a scalar
epilogue/remainder of the loop might be used to handle the tail iterations. (3) Various runtime
checks might be inserted throughout the vectorized control flow to handle, for example, pointers 
to overlapping memory regions, in which case the scalar version of the loop is executed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{LLVM's Vectorization Plan}
\label{sec:vplan}

\subsection{History}
LLVM implements two different vectorizers for auto-vectorization, LoopVectorize (LV)
and SLPVectorize. These vectorizers are a part of LLVM's middle-end and run after module 
simplification has taken place. As VPlan still needs to be integrated into SLP-Vectorization, we 
will only focus on explaining the original LoopVectorize structure.

LoopVectorize was designed to focus only on vectorization of innermost loops~\cite{llvmintrvplan}. 
When starting with an initial scalable loop in IR, it would first run a legality phase to check 
whether  and how it is legal to vectorize the loop. This phase would produce artifacts/maps, such as if 
and what runtime aliasing checks would be required. Then, the cost model ran next, creating more 
artifacts based on the scalar IR and legality artifacts. These cost model artifacts would 
associate different vectorization factors with a cost. Lastly, the transformation phase would vectorize 
the underlying IR in one go based on all the produced artifacts.

However, this rigid flow of LoopVectorize presented a challenge. Extending LV to include outer loop 
and nested loop vectorizations would have been difficult, as LV was designed to flatten a
loop's control flow into a single basic block. Producing and considering more and 
more artifacts when trying to transform the scalar loop into a vectorized one all at once also did 
not scale well.

As a potential solution, Intel proposed a long-term effort to refactor LLVM's whole auto-vectorization
using a system called Vectorization Plan (VPlan)~\cite{llvmextloopvec}.

The goal was to unite any auto-vectorization systems (SLP, Inner-/Outer-Loop) into a single flexible
system~\cite{llvmintrvplan,llvmvplanstate}, as seen in Figure~\ref{fig:vplan-future}. This system would 
use VPlans to represent different 
vectorization approaches abstracted away from the underlying IR. Each VPlan would 
calculate its cost and transform itself to modify the final IR. At first, the Legality phase (yellow) 
would check if the scalar code would be legal to vectorize at all, and if so, initial VPlans would 
be constructed. In the following Planning phase (magenta), these initial VPlans would be optimized 
using various VPlan-to-VPlan transformations~\cite{llvmouterloopstatus}. 
The VPlan's cost model would supervise those transformations. Finally, the best VPlan would be chosen 
and materialized/executed, modifying the underlying IR.

\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/vplan-future.png}
  \caption{The planned future architecture for VPlan. The VPlan optimization would
  work as iteratively applying VPlan-to-VPlan transformations.}
  \label{fig:vplan-future}
\end{figure}

\subsection{State of VPlan (LLVM 19.1.4)}
\emergencystretch 6em

In its current state, VPlan is deeply integrated into LV.
When auto-vectorizing, LV starts by checking all the different loops in the IR.
If it finds one, it will try to vectorize it (\texttt{LoopVectorizePass::\allowbreak processLoop}).
Several factors might hinder vectorization for a given loop: 
(1) If the loop was already vectorized before, it will not be vectorized again. 
(2) The Legality phase checks if and how it is legal to vectorize the scalar
loop. (3) The loop's trip count might be too small to make vectorization profitable. (4) The cost 
of runtime checks, such as pointer aliasing, can be too high.

The initial VPlans are constructed based on the legality checks and the legacy LV cost model's 
cost decisions referencing the scalar IR (\texttt{LoopVectorizationPlanner::\allowbreak plan}). 
After initial VPlan construction, various VPlan-to-VPlan 
transformations are applied to optimize it 
(\texttt{LoopVectorizationPlanner::\allowbreak buildVPlansWithVPRecipes}).

Then, the optimal vectorization factor (VF) is calculated based on the set of VPlans and their 
representative cost (\texttt{LoopVectorizationPlanner::\allowbreak cost}), the calculation of 
which still involves the legacy LV cost model.
The best VPlan is picked based on the optimal VF. A VF of 1 means it is not profitable to vectorize 
the loop; however, interleaving/unrolling can still modify the underlying IR. Afterwards, the 
Interleave Count (IC) is determined 
(\texttt{LoopVectorizationCostModel::\allowbreak selectInterleaveCount}).

The best VPlan is then executed (\texttt{LoopVectorizationPlanner::\allowbreak executePlan}). In the 
case of vectorization (VF $>1$), LV will also directly try to vectorize the epilogue loop.

\subsubsection{VPlan Structure}
To better understand the structure of a VPlan, we will use the
simple conditional loop from Figure~\ref{fig:inner-loop-vec} as an example.

Figure~\ref{fig:inner-loop-scalar-ir} shows the scalar IR of the main loop basic blocks \texttt{for.body},
\texttt{if.then}, and \texttt{for.inc}. Figure~\ref{fig:inner-loop-vplan-hcfg-body} shows the equivalent
hierarchical control-flow-graph (CFG) of a VPlan modeling a vectorization approach for this 
scalar loop. Other basic blocks like the scalar epilogue loop and its predication check (is the final
vector trip count equal to the original trip count) are omitted. See Figure~\ref{fig:inner-loop-vplan-hcfg}
in Appendix~\ref{appendix-figures} for the fully modeled HCFG.

%TODO: Combine inner-loop-scalar-ir figure and inner-loop-vplan-hcfg into a single
%       double column figure.

\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/inner-loop-scalar-loop-ir-color.png}
  \caption{The scalar IR for the main loop part of the conditional loop depicted in
  Figure~\ref{fig:inner-loop-vec}. Note that \texttt{getelementptr inbounds}
  was abbreviated to \texttt{getelmptr inbs}.}
  \label{fig:inner-loop-scalar-ir}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/inner-loop-vplan-hcfg-loop-body-color.png}
  \caption{The vectorized loop part of a hierarchical CFG of a VPlan modelling the conditional loop
  depicted in Figure~\ref{fig:inner-loop-vec}. Some predefined values are
  \texttt{vp<\%0>} = VF * UF; \texttt{vp<\%1>} = vector-trip-count; \texttt{vp<\%2>} = original trip-count.}
  \label{fig:inner-loop-vplan-hcfg-body}
\end{figure}

The different colored parts in the scalar IR can be found with the same color in the modeled VPlan.
The only completely new part that VPlan models in this example is the handling of the scalar epilogue
loop (purple).

The hierarchical CFG of a VPlan helps abstract it away from the underlying IR. The nodes of this CFG
are either \texttt{VPBasicBlocks}, containing a sequence of one or more \texttt{VPRecipes}, or
\texttt{VPRegionBlocks}, modelling a single-entry-single-exit subgraph.
In our example the \texttt{vector.body} is a basic block and \texttt{vector loop} is a region block.

\texttt{VPRegionBlocks} can indicate that their contents are to be replicated a certain number of times,
to model more complex control flow for different VF, UF combinations more efficiently in a single VPlan. 
In our example \texttt{vector loop} is not to be replicated (\texttt{<x1>}).

Inside our two \texttt{VPBasicBlocks} we have a list of recipes (lines of instructions). A recipe
models one or more instructions in the executed IR and can be based on one or more instructions.
from the input IR (ingredients). Note that a recipe might also model a instruction with no reference to any input
IR (in our example this would be the inversion of our vector mask). There are different types of 
recipes. \texttt{VPReplicateRecipe} will either \texttt{CLONE} exactly one instruction from the input IR
into exactly one instruction in the output IR, or \texttt{REPLICATE} an instruction a certain number
of times into output IR (typically found in replicator \texttt{VPRegionBlocks}). Different 
\texttt{VPWidenRecipes} exist to transform/\texttt{WIDEN} any input IR instruction into a vectorized 
version of it. The last remaining recipe type in our example are \texttt{VPInstructions}. They
\texttt{EMIT} new instructions into the output IR that have no representative in the input IR (no
ingredients).

Furthermore, VPlan recipes can also model definitions and usages of different values. In our example,
the comparison mask \texttt{vp<\%6>} is a \texttt{VPValue} that is used by three different 
\texttt{VPUsers}.

\subsection{Future of VPlan}
Many of the VPlan internals and systems surrounding it are still in active development. To stay
up-to-date with the latest developments regarding auto-vectorization in LLVM one can follow the 
Community Forum\footnote{\url{https://discourse.llvm.org/c/ir-optimizations/loop-optimizations/62}} or
attend the monthly online 
meetings\footnote{\url{https://discourse.llvm.org/t/monthly-vectorizer-online-sync-up/78978/3}}.

\subsubsection{VPlan and Integration with LV}
The main effort of setting up the VPlan structure and integrating it first into LV has seen
tremendous effort over the last years. Florian Hahn presented future directions for 
VPlan~\cite{llvmvplanupdate}, which mostly evolve around transforming more LV-based decisions into
explicit VPlan-to-VPlan transformations. Additionally, a lot of cost model related decisions 
still end up being influenced by the old legacy LV cost model, however, transition to a VPlan-based 
cost model has already started.

\subsubsection{Outer Loop Vectorization}
Development around outer loop vectorization has slowed down a bit as focus shifted more onto VPlan
internals and integration into LV. The inner loop path and outer loop path are currently still
seperated by the experimental \texttt{-enable-\allowbreak vplan-\allowbreak native-\allowbreak path} 
flag. Further developing the 
outer loop path and eventually merging it with the innermost loop path is planned.

\subsubsection{Merging with SLP Vectorize}
Merging the existing SLP Vectorizer with the current VPlan infrastructure is in the early phase
of planning.

\subsubsection{Function Vectorization}
As far as we now there is currently no active development for whole-function vectorization in the
llvm-project's upstream yet.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{sec:relatedwork}
Besides the main llvm-project's upstream there are many different projects making use of or extending
LLVM's toolchain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary and Future Work}
\label{sec:summary}
In conclusion, VPlan has been in active development for several years now. When compiling with
LLVM VPlan is used to model initial vectorization approaches, keep track of different
optimization decisions and finally execute the best VPlan to transform IR. 
However, there are still some VPlan-to-VPlan transformations and a VPlan-based cost model missing, 
before the legacy LV code can be completely discarded.

Outer loop vectorization support via VPlan is still experimental, seperated from the inner loop
LV integration, and runs on a divergent experimental path (\texttt{-enable-\allowbreak vplan-\allowbreak native-\allowbreak path}).
In the near future development on outer loop vectorization will continue and merging it with LV's
VPlan path is planned.

Active development of other approaches such as SLP and whole-function vectorization based on the 
VPlan architecture has not been started yet.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper} % read paper.bib file

\appendix

\section{Appendix: Additional Figures}\label{appendix-figures}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{images/inner-loop-vplan-hcfg.png}
  \caption{Full hierarchical CFG of a VPlan modelling the vectorized version of the loop 
  depicted in Figure~\ref{fig:inner-loop-vec} as of LLVM 19.1.4.}
  \label{fig:inner-loop-vplan-hcfg}
\end{figure*}

\end{document}